Community biomass and species richness temporal trends reveal contrasting assembly and disassembly processes in natural food-webs

### 2022-02-23 paper/manuscript.Rmd
Check introduction, matmet and discussion

---
title: "Community biomass and species richness temporal trends reveal contrasting
    assembly and disassembly processes in natural food-webs"
author: Alain Danet, Loubna El Madouri, Maud Mouchet, Elisa Thébault and Colin Fontaine (ordre non fixé)
date: \today 
output:
  bookdown::pdf_document2:
    fig_caption: true 
    keep_tex: true
    toc: true
  bookdown::word_document2:
fontsize: 12pt
header-includes:
   - \usepackage{natbib}
   - \usepackage{float}
   - \usepackage[table]{xcolor}
   - \usepackage{multirow}
   - \usepackage{booktabs}
   - \usepackage{makecell}
   - \usepackage{setspace}
   - \doublespacing
   - \usepackage{lineno}
   - \linenumbers
geometry: margin=2cm
bibliography: references.bib
number_sections: true
---

@@

Multiple studies reported dramatic declines of population and community biomass
[@ceballos_biological_2017; @hallmann_more_2017; @albano_native_2021], as well
as decrease in species richness [@barnosky_has_2011; @newbold_global_2015] and
turnover in species composition [@comte_climate_2021]. Unfortunately, studies often
focus on the temporal trend of one dimension of biodiversity [@hallmann_more_2017,
etc, but see @]. Because the functioning of ecosystem depends ultimately on
multiple facets of biodiversity, such as species richness, evenness and
functional composition^[1]  [REF @thibaut_understanding_2013], the documentation
of the trend of one biodiversity facet is likely insufficient to evaluate
precisely its consequences
on ecosystem functioning.

@@

Current knowledge on the links among different facets of biodiversity mainly
comes from from spatial comparison [@dunne_network_2006;
@soininen_relationship_2012; @grace_integrative_2016; @danet_species_nodate]. space-for-time substitution began to be widely
used in the early 20th century to infer the temporal succession of vegetation
from space variations [@pickett_space-for-time_1989]. It supposes a homology
between variations in space and in time, i.e. that spatial sequences correspond
to temporal sequences.  This method have raised criticisms
[@damgaard_critique_2019], with studies showing analogous patterns in space and
time [@blois_space_2013] or not [@johnson_testing_2008;
@adler_contrasting_2007]. Spatial turnover of species composition, dispersal
limitation, disturbance and stochastic processes have been hypothesised to be
responsible for the non-compliance of space-for-time substitution
[@adler_contrasting_2007; @blois_space_2013].

@@

The documentation of biodiversity temporal trends generally neglect species
interactions although they are fundamentals to ecosystem functioning
[@thebault_food-web_2003; @oconnor_general_2017]. In food-webs, spatial
comparisons in empirical and theoretical studies generally reported that trophic
level positively scales to community biomass [@wang_biodiversity_2018;
@danet_species_nodate]. Trophic level has also been shown to stabilise the
temporal variation of community biomass [@shanafelt_stability_2018;
@danet_species_nodate]. Connectance, which measures the degree of generalism in
a network, was shown to increase robustness to extinction [@dunne_network_2002].
However, the temporal variation of food-web structure is largely unknown
[@kortsch_disentangling_2021]. Food-web temporal variation are often reported
across small time-steps [@warren_spatial_1989; @pecuchet_novel_2020;
@kaartinen_high_2012]. Our knowledge are even more scarce on the variability of
food-web structure in time and particularly how those temporal variations are linked to
the temporal variations of other aspects of biodiversity such as species richness
and community biomass. Spatially, connectance has been shown to be negatively
related to species richness in empirical food-webs [reviewed in
@dunne_network_2006]. Connectance has been shown to scale positively or
negatively with community biomass in theoretical models
[@thebault_food-web_2003; @poisot_trophic_2013] (*Add mechanism?*). In order to
improve our understanding on the assembly mechanisms in food-webs, it is thus
crucial to assess how biomass and species richness temporal trends are associated to
temporal trends in food-web structure.

In this line, how the temporal changes of community biomass are related to
temporal changes of species richness is still questionned in natural settings,
especially in food-webs. Our knowledge on the relationships between food-web
structure and other facets of biodiversity often comes from the comparison of
food-web snapshots along spatial gradient, raising the question of the validity
of space-for-time substitution to infer consequences of species richness and
community biomass variations on food-web structure.

The increasing availability ecological temporal series and the development of
methods to reconstruct food-webs [@bonnaffe_comparison_2021;
@gravel_inferring_2013; @danet_species_nodate] makes possible to assess those
questions. We used the standardized monitoring of fish freshwater communities on
the period 1995-2018 over France, documenting the individual size of fishes in
400 sites during ten years or more. We then reconstruct around 4000 food-webs
based on allometric relationships and food diet data. We questionned i) how
temporal trends of food-web structure are  linked to trends of community biomass
and trends in species richness? ii) Are the relationships observed in time the
same as those observed in space, i.e. is the space-for-time substitution valid?

@@

We selected the sites that have been monitored at least 10 years on the
1995-2018 period, with the same sampling protocol over time. We selected
one sampling a year by site. In each site, we ensured low temporal heterogeneity
in sampling period by selecting sampling that were performed in the same period,
leading to a mean standard deviation of sampled month in each site of three
weeks [@danet_species_nodate]. Those steps leads to the selection of 
`r nb_st` sites (Figure \@ref{fig:matmet}A).

@@

We first inferred the body mass of each fish individual considering an allometric
law, $B_i = 0.01 \times l_i^3.03$, $B_i$ being the biomass expressed in gram and $l_i$ the
body length of the individual $i$ expressed in millimetres  [@froese_cube_2006]. Then the
biomass of each species and community biomass were obtained by summing the body
mass of individuals. Finally, the biomass was reported to the sampled surface
(in $m^2$) to account for sampling effort.

@@

We inferred food-web structure in two steps [@bonnaffe_comparison_2021;
@danet_species_nodate]: (1) we first computed the metaweb, describing the
potential trophic interactions among all size classes species, (2) then we
extracted each yearly snapshot of food-webs by extracting species size
classes and their interactions according to their presence in a given sampling
event.

We inferred the metaweb according to ontogenic diet data and body size
[@gravel_inferring_2013; @poisot_structure_2016; @brose_predator_2019;
@bonnaffe_comparison_2021; @danet_species_nodate]. The method is extensively
described in @bonnaffe_comparison_2021 and @danet_species_nodate. The food diet
of stream fishes changes along their life cycle or stage, therefore
according to their body size. Each fish species was divided in nine size
classes, named "trophic species", equally distributed in their observed body size
range. Thus, a trophic species is a group of individuals of the same species and size
class that are expected to share the same preys and predators. We also added
seven nodes present in the ontogenic food diet database: detritus, biofilm,
phytoplankton, zooplankton, macrophages, phytobenthos and zoobenthos.  The
resulting metaweb contained 45 fish species divided into nine size classes plus
7 resource nodes, summing up to 412 nodes.  Consideration of size classes, i.e.
trophic species instead of species, improves inference of trophic interactions
[@jennings_weak_2001].

@@

To assess the validity of the space-for-time substitution hypothesis, we
compared the relationships among food-web structure in time to the ones observed
along spatial gradient. To do so, we selected the same sites that were used
for the temporal analysis. We summarised temporal series of each station by the
median value over time of each facet of food-web structure described above.

@@

We modelled the relationships among the temporal trends of food-web structure ($dN$)
depending on the temporal trends of community biomass ($dB$) and species
richness ($dR$) with a linear model. in order to test the relationship among the temporal of food-web
structure, that is connectance, average trophic level and evenness of biomass
from one side and species richness and community biomass from the other side as
follows:

@@

In order to control for the uncertainties of the estimated temporal trends, each
temporal trend was weighted by the inverse of the product of the standard errors of the
estimated trends, i.e. slopes for the response and the explicative
variables ($w_j = \dfrac{1}{s_{r_j} s_{e,j}}$, $w_j$ being the weight, $s_{r_j}$ and $s_{e_j}$ being
respectively the standard errors on the temporal trends of the response and
explicative variables). We also included a statistical model including the
interaction between the temporal trends of species richness and community
biomass (i.e.
adding $\beta_3dB_jdR_j$), the results are presented in supplementary material (Fig.
SXX). To model the relationships between the temporal trends of community
biomass and the one of species richness, we built the model as $dB_j = \alpha_j +
\beta_1dR_j + \epsilon_j$.

@@

Spatial gradients of food-web structure are partly driven by environmental gradients, so
we controlled relationships among variables by adding two PCA axis
synthesizing respectively (1) stream size and (2) average temperature and
enrichment. Enrichment was characterized by the average Biological Oxygen
Demand [BOD, @mallin_factors_2006][^2].

@@

  - $N_j$: median network structure and communities structure
  - $B_j$: median community biomass
  - $R_j$: median species richness 
  - $E_{1,j}$: PCA1 related to stream size
  - $E_{2,j}$: PCA2 related to average temperature and average BOD
  - $\epsilon_{\alpha|i}$: random effect of the hydrographic basin
    $i$ on $\alpha$

@@

The significativity of the effects were tested with ANOVA. In order to compare
the magnitude of the effects of community biomass and species richness in both
temporal and spatial models, we computed standardized slopes ($r_{\delta}$) and
their associated standard errors as $r_{\delta} = \beta\frac{s_x}{s_y}$.
With $\beta$ being the unstandardised slope, $s_x$ and $s_y$ the standard
deviation of the explicative and response variables.

All the statistical models displayed low multicollinearity (VIF < 3, Table XX).
We visually inspected the residuals of the linear models (TODO).

@@

The temporal trends of connectance were positively linked with the ones of
community biomass ($r_\delta = `r tps_bm_sest["ct"]`$), also meaning that both
temporal trends had the same direction in average. That is, the sites where
food-web connectance decreased over time also had community biomass that
decreased over time and reversely. It suggests that communities that loose biomass
over time, also loose generalist predators and reversely that communities
gaining biomass over time also gain generalist predators. Spatially, we also found
a positive relationship between median connectance and median species
richness ($r_\delta = `r sp_b_sest["ct"]`$). However, we did not observe
significant relationships between the temporal trends of connectance and the
ones of species richness (Fig. \@ref(fig:bmr)D) while we found a stong
negative relationship between median connectance and median species richness
along the spatial gradient
($r_\delta = `r sp_r_sest["ct"]`$, Fig. \@ref(fig:std)).

We found that the temporal trends of average trophic level had a strong positive
relationship with the ones of community biomass 
($r_\delta = `r tps_bm_sest["tlvl"]`$, Fig. \@ref(fig:bmr)B, Table
\@ref(tab:aov)), meaning
that sites where community biomass decreased over time, average trophic level
decreased over time and *vice versa*. It suggests that in communities loosing biomass
over time, the average biomass of the food-web moved down to lower trophic level
and *vice versa*. The spatial gradient also showed a positive relationship between
median average trophic level and median community biomass, but half the strength 
as the temporal one ($r_\delta = `r sp_b_sest["tlvl"]`$, Fig. \@ref(fig:std),
Table \@ref(tab:aov-sp)). The temporal trends of average
trophic level was unrelated to the ones of species richness. It suggests that in
communities either loosing or gaining species richness over time, the location of
average biomass along trophic level did not significantly change according to species richnnes.  However, we
found a strong negative relationship in space 
($r_\delta = `r sp_r_sest["tlvl"]`$, Fig. \@ref(fig:std), Table
\@ref(tab:aov-sp)).

@@

```{r bmr, out.width="70%", fig.height=7, fig.cap=fig_bm_rich_cap}
plot_grid(
  plot_grid(plotlist = fig2$bm,
    ncol = 1, labels = c("A", "B", "C"),
    hjust = -5.5, vjust = 1.75
    ),
  plot_grid(plotlist = fig2$rich,
    ncol = 1, labels = c("D", "E", "F"),
    hjust = -2, vjust = 1.75
    ),
  ncol = 2,
  rel_widths = c(1, 1)
)
```

@@

We further found that most temporal trends of food-webs were classified as
stable for species richness and community biomass. We found almost the same
number of stations that displayed temporal decrease or increase for species
richness, and twice more stations displayed decrease in community biomass than
increase. This result is in accordance with previous meta analysis on species
richness trends of vertebrate in marine, freshwater and terrestrial systems
[@dornelas_assemblage_2014; @blowes_geography_2019; @klink_meta-analysis_2020].
While the spatial coverage of our data is good, it would be rather risky to
conclude about biodiversity trends in freshwater fish communities because we did
not assess pressures that may affect the temporal trends such as land-use
changes and species invasions [@comte_climate_2021; @su_human_2021;
@cardinale_is_2018].

@@

The temporal trends of connectance were positively linked to that of community
biomass, meaning that when connectance increased through time when community
biomass increased through time in a given site. Reversely, connectance decreased
in food-webs when community biomass decreased through time. It suggests that
community biomass losses and gains are associated to losses and gains in large
bodied piscivores. This interpretation is in
line with the results that community biomass losses through time were associated
to a temporal decrease of average trophic level and that community biomass
temporal gains were associated to temporal increase of average trophic level.
Another piece of evidence is that evenness of biomass, measured by the Pielou
index, decreased through time as community biomass increased, meaning that
biomass distribution across species becomes less even when community biomass
increased. Reversely, in food-webs that displayed temporal decrease in community
biomass, we found that evenness of biomass increased, i.e. that biomass
distribution across species becomes more even. Because predator-prey
interactions relate in part on bodysize and thus bodymass to a certain extent,
we may expect that a given gain in predators of higher trophic level leads to a
decrease of biomass evenness and reversely. In the same line, we found that the
temporal trends of community biomass were positively accociated to the temporal
trends of the number and proportion of piscivorous trophic species present in
fish communities (Fig. SXX).

In contrast, we found that temporal trends of connectance and average trophic
level were not related to the temporal trends of species richness. However, we
found that temporal trends of the number of piscivorous species and piscivorous
trophic species were positively associated to the temporal trends of species
richness (Fig. S3). It suggests that the temporal decrease of species richness
observed in our data are concomimant to losses of piscivorous and
non-piscivorous species, leading to the absence of temporal trends in
connectance and average trophic level and symmetrically for the temporal
increase of species richness. Furthemore, we found that the temporal trends of
biomass evenness had a positive relationship with the one of species richness,
in contradiction of what we found for temporal trends of community biomass.
Together with the fact that species richness temporal trends had a positive
relationship with the temporal trends of community biomass, it suggests that
positive and negative  temporal trends of species richness are linked to gains
and losses of redundancy in food-webs, without overall effects on connectance
and average trophic level. This is in line with the insurance hypothesis
[@yachi_biodiversity_1999] which predicts the temporal loss and gain of species
richness represent respectively loss and gain of redundancy in the community,
here in the food-web structure.

As studies reported community biomass or species richness temporal trends
[@hallmann_more_2017; @ceballos_biological_2017; @dornelas_assemblage_2014;
@blowes_geography_2019], we document that temporal trends of species richness
and community biomass might have different consequences on the structure of the
community. In particular, the decrease of the average trophic level associated
to the loss of community biomass could have cascading effect on ecosystem
functioning as the highest predators are predicted to increase productivity and
stabilise communities [@wang_biodiversity_2018; @shanafelt_stability_2018;
@danet_species_nodate]. In contrasts, the loss of species richness might be
linked to loss of redundancy in the food-web, highlighting that those
communities may become more at risk regarding future perturbations
[@yachi_biodiversity_1999]. Overall, it suggests that monitoring several facets
of community structure is essential to infer how ecological communities are
reshaped by loss and gain of biomass and species richness.

@@

Surprisingly, we found that the temporal trends of connectance were not related
to the one of species richness as found spatially in most empirical food-webs
[@dunne_network_2006] and also here. We explained this pattern by the fact that
the temporal variation of species richness in our dataset seems to be linked to
temporal variation of redundancy level in the structure of food-webs rather than
overall change of their vertical structure. The same rational can applied with
average trophic level. This result is in line with other studies which found
mismatch between spatial and temporal patterns [@adler_contrasting_2007;
@johnson_testing_2008]. This discrepency in space and time might be due to the
crucial assumption that variations in space reflect differences in assembly
stages of ecological communities and then mirrors their temporal
assembly/disassembly trajectories. Temporal variation of species richness might
relate on transitory dynamics while spatial gradient reflect aggregated dynamics
over longer time scale including a far number of processes including high
species turnover along spatial gradient. However, we fount that the positive
relationship between evenness of biomass and species richness were consistent in
space and time. Another reason might be that the spatial gradient displays a
higher range of variation for species richness than the temporal stations (Fig.
S2). However, we found strong relationships between the temporal trend of
species richness from one side and the temporal trends of community biomass,
biomass evenness, and the number of piscivore species from the other side (Fig. S3).

>>>>

---
title: "Community biomass and species richness temporal trends reveal contrasting assembly and disassembly processes in natural food-webs"
author: Alain Danet, Loubna El Madouri, Maud Mouchet, Elisa Thébault and Colin Fontaine
date: \today
output:
  bookdown::pdf_document2:
    fig_caption: true 
    keep_tex: true
    toc: true
  bookdown::word_document2:
fontsize: 12pt
header-includes:
   - \usepackage{natbib}
   - \usepackage{float}
   - \usepackage[table]{xcolor}
   - \usepackage{multirow}
   - \usepackage{booktabs}
   - \usepackage{makecell}
   - \usepackage{setspace}
   - \doublespacing
   - \usepackage{lineno}
   - \linenumbers
geometry: margin=2cm
bibliography: references.bib
number_sections: true
---

@@

Multiple studies reported dramatic declines of population and community biomass
[@ceballos_biological_2017; @hallmann_more_2017; @albano_native_2021], as well
as decrease in species richness [@barnosky_has_2011; @newbold_global_2015] but
also recent studies reported non significant temporal trends in species richness
[@vellend_global_2013; @dornelas_assemblage_2014; @blowes_geography_2019], total
abundance [@klink_meta-analysis_2020], high turnover in species composition
[@dornelas_assemblage_2014; @blowes_geography_2019]. Unfortunately, studies
often focus on the temporal trend of one dimension of biodiversity
[@hallmann_more_2017; @klink_meta-analysis_2020, but see 2 in
@dornelas_assemblage_2014; @blowes_geography_2019]. However biodiversity is a
multidimensional concept and the functioning of ecosystem depends ultimately on
several dimensions of biodiversity, such as species richness, evenness and
functional composition^[1]  [@cardinale_biodiversity_2012; @duffy_why_2009], the
documentation of the trend of one biodiversity facet is likely insufficient to
understand changes in community structure and ultimately predict their
consequences on ecosystem functioning.

@@

Current knowledge on the links among the different facets of biodiversity mainly
comes from spatial comparison [@dunne_network_2006;
@soininen_relationship_2012; @grace_integrative_2016; @danet_species_nodate],
which is used to predict how community structure will change in time, under the
assumption of the space-for-time substitution. It supposes a homology between
variations in space and in time, i.e. that spatial sequences correspond
to temporal sequences. This method have raised criticisms
[@damgaard_critique_2019], with studies showing analogous patterns in space and
time [@blois_space_2013] or not [@johnson_testing_2008;
@adler_contrasting_2007]. Spatial turnover of species composition, dispersal
limitation, disturbance and stochastic processes have been hypothesised to be
responsible for the non-compliance of space-for-time substitution
[@adler_contrasting_2007; @blois_space_2013]. The recent increase in the availability
of temporal monitoring of ecological communities makes possible to address
biodiversity trends for multiple facets.

@@

More dramatic is that the temporal trends of biodiversity generally neglect
species interactions. Although species interactions determine for a large part
community composition, species coexistence and ecosystem functioning
[@chesson_mechanisms_2000; @thebault_food-web_2003; @oconnor_general_2017].
Furthermore, species interactions can imply tremendous secondary extinctions
[@paine_note_1969; @dunne_network_2002]. In food-webs, spatial comparisons in
empirical and theoretical studies generally reported that trophic level
positively scales to community biomass [@wang_biodiversity_2018;
@danet_species_nodate]. Trophic level has also been shown to stabilise the
temporal variation of community biomass [@shanafelt_stability_2018;
@danet_species_nodate]. Connectance, which measures the degree of generalism in
a network, was shown to increase robustness to extinction [@dunne_network_2002].
However, the temporal variation of food-web structure is largely unknown
[@kortsch_disentangling_2021]. Food-web temporal variation are often reported
across small time-steps [@warren_spatial_1989; @pecuchet_novel_2020;
@kaartinen_high_2012]. Our knowledge on the variability of food-web structure in
time are even more scarce and particularly how those temporal variations are
linked to the temporal variations of other aspects of biodiversity such as
species richness and community biomass. Spatially, connectance has been shown to
be negatively related to species richness in empirical food-webs [reviewed in
@dunne_network_2006]. Connectance has been shown to scale positively or
negatively with community biomass in theoretical models
[@thebault_food-web_2003; @poisot_trophic_2013]. Average trophic level was
previously shown to scale negatively with species richness
[@danet_species_2021].  In order to improve our knowledge on the consequences of
biodiversity trends such as community biomass and species richness, it is thus
crucial to assess how biomass and species richness temporal trends are
associated to temporal trends of food-web structure.

Even how the temporal trends of community biomass are related to temporal
trends of species richness is still questionned in natural settings, especially
in food-webs. Our knowledge on the relationships between food-web structure and
other facets of biodiversity often comes from the comparison of food-web
snapshots along spatial gradient, raising the question of the validity of
space-for-time substitution to infer consequences of species richness and
community biomass variations on food-web structure.

The increasing availability ecological temporal series and the development of
methods to reconstruct food-webs [@bonnaffe_comparison_2021;
@gravel_inferring_2013; @danet_species_nodate] makes possible to assess the
relationships among several biodiversity facets, including food-web structure.
We used the standardized monitoring of fish freshwater communities on the period
1995-2018 over France, documenting the individual size of fishes in
400 sites during ten years or more. We then reconstruct around 4000 food-webs
based on allometric relationships and food diet data. We questionned i) how
temporal trends of food-web structure are  linked to trends of community biomass
and trends in species richness? ii) Are the relationships observed in time the
same as those observed in space, i.e. is the space-for-time substitution valid?

@@

We selected the sites that have been monitored at least 10 years on the
1995-2018 period, with the same sampling protocol over time. We selected
one sampling a year by site. In each site, we ensured low temporal heterogeneity
in sampling period by selecting sampling that were performed in the same period,
leading to a mean standard deviation of sampled month in each site of three
weeks [@danet_species_nodate]. Those steps leads to the selection of 
`r nb_st` sites (Figure \@ref(fig:matmet)A).

@@

We first inferred the body mass of each fish individual considering an allometric
law, $B_i = 0.01 \times l_i^{3.03}$, $B_i$ being the biomass expressed in gram and $l_i$ the
body length of the individual $i$ expressed in millimetres  [@froese_cube_2006]. Then the
biomass of each species and community biomass were obtained by summing the body
mass of individuals. Finally, the biomass was reported to the sampled surface
(in $m^2$) to account for sampling effort.

@@

We inferred food-web structure in two steps [@bonnaffe_comparison_2021;
@danet_species_2021]: (1) we first computed the metaweb, describing the
potential trophic interactions among all size classes species, (2) then we
extracted each yearly snapshot of food-webs by extracting species size classes
and their interactions according to their presence in a given sampling event.

We inferred the metaweb according to ontogenic diet data and body size
[@gravel_inferring_2013; @poisot_structure_2016; @brose_predator_2019;
@bonnaffe_comparison_2021; @danet_species_2021]. The method is extensively
described in @bonnaffe_comparison_2021 and @danet_species_2021. The food diet of
stream fishes changes along their life cycle or stage, therefore according to
their body size. Each fish species was divided in nine size classes, named
"trophic species", equally distributed in their observed body size range. Thus,
a trophic species is a group of individuals of the same species and size class
that are expected to share the same preys and predators because trophic
interactions are largely determined by predator-prey size ratio
[@brose_allometric_2006]. We also added seven nodes present in the ontogenic
food diet database: detritus, biofilm, phytoplankton, zooplankton, macrophages,
phytobenthos and zoobenthos.  The resulting metaweb contained 45 fish species
divided into nine size classes plus 7 resource nodes, summing up to 412 nodes.
Consideration of size classes, i.e.  trophic species instead of species,
improves inference of trophic interactions [@jennings_weak_2001].

@@

To assess the validity of the space-for-time substitution hypothesis, we
compared the relationships among food-web structure in time to the ones observed
in space. To do so, we selected the same sites that were used
for the temporal analysis. We summarised temporal series of each station by the
median value over time of each facet of food-web structure described above.

@@

We modelled the relationships among the temporal trends of food-web structure
($dN$), that is connectance, average trophic level and evenness of biomass, and
the temporal trends of community biomass ($dB$) and species richness ($dR$) with
Ordinary Least Squares. The simultaneous assessment of the
effects of species richness and community biomass allows to tease apart their
respective effect on food-web structure.

@@

In order to control for the uncertainties of the estimated temporal trends, each
temporal trend was weighted by the inverse of the product of the standard errors
of the estimated trends, i.e. estimated slopes for the response and the
explicative variables ($w_j = \dfrac{1}{s_{r_j} s_{e,j}}$, $w_j$ being the
weight, $s_{r_j}$ and $s_{e_j}$ being respectively the standard errors on the
temporal trends of the response and explicative variables). We also included a
statistical model including the interaction between the temporal trends of
species richness and community biomass (i.e.  adding $\beta_3dB_jdR_j$), the
results are presented in supplementary material (Fig.  SXX). To model the
relationships between the temporal trends of community biomass and the one of
species richness, we built the model as $dB_j = \alpha_j + \beta_1dR_j +
\epsilon_j$.

@@

Spatial variation of food-web structure are partly driven by environmental gradients, so
we controlled relationships among variables by adding two PCA axis
synthesizing respectively (1) stream gradient from upstream to downstream and
(2) enrichment (Fig. \@ref(fig:matmet)). Enrichment was characterized by the
average Biological Oxygen Demand [BOD, @mallin_factors_2006][^2]. We further
added a random effect of the hydrographic basin on the intercept.

@@

  - $N_j$: median food-web structure 
  - $B_j$: median community biomass ($g.m^-2$) 
  - $R_j$: median species richness ($m^-2$) 
  - $E_{1,j}$: First PCA axis 
  - $E_{2,j}$: Second PCA axis
  - $\epsilon_{\alpha|i}$: random effect of the hydrographic basin
    $i$ on $\alpha$

@@

In order to compare the magnitude of the effects of community biomass and
species richness in both temporal and spatial models, we computed standardized
slopes ($r_{\delta}$) and their associated standard errors as $r_{\delta} =
\beta\frac{s_x}{s_y}$.  With $\beta$ being the unstandardised slope, $s_x$ and
$s_y$ the standard deviation of the explicative and response variables.

All the statistical models displayed low multicollinearity (Variance inflation
factor < 3, Table XX).  We visually inspected the residuals of the linear models
(TODO).

@@

The temporal trends of connectance were positively linked with the ones of
community biomass ($r_\delta = `r tps_bm_sest["ct"]`$), also meaning that both
temporal trends had the same direction in average. That is, the sites where
food-web connectance decreased over time also had community biomass that
decreased over time and reversely. It suggests that communities that decreased
in biomass
over time loose generalist predators over time, as connectance measures the degree of generalism and reversely that communities
gaining biomass over time also gain generalist predators. Spatially, we also found
a positive relationship between median connectance and median species
richness ($r_\delta = `r sp_b_sest["ct"]`$). However, we did not observe
significant relationships between the temporal trends of connectance and the
ones of species richness (Fig. \@ref(fig:bmr)D) while we found a stong
negative relationship between median connectance and median species richness
along the spatial gradient
($r_\delta = `r sp_r_sest["ct"]`$, Fig. \@ref(fig:std)).

The temporal trends of average trophic level had a strong positive
relationship with the ones of community biomass 
($r_\delta = `r tps_bm_sest["tlvl"]`$, Fig. \@ref(fig:bmr)B, Table
\@ref(tab:aov)), meaning
that sites where community biomass decreased over time, average trophic level
decreased over time and *vice versa*. It means that in communities loosing biomass
over time, the average biomass of the food-web moved down to lower trophic level
and *vice versa*. We also found a positive relationship between
median average trophic level and median community biomass in space, but half the strength 
as the temporal one ($r_\delta = `r sp_b_sest["tlvl"]`$, Fig. \@ref(fig:std)). The temporal trends of average
trophic level was unrelated to the ones of species richness. It suggests that in
communities either loosing or gaining species richness over time, the location of
average biomass along trophic level did not significantly change according to
species richnnes.  However, we found a strong negative relationship in space 
($r_\delta = `r sp_r_sest["tlvl"]`$, Fig. \@ref(fig:std), Table
\@ref(tab:aov-sp)).

@@

```{r bmr, out.width="80%", fig.height=7, fig.cap=fig_bm_rich_cap}
plot_grid(
  plot_grid(plotlist = fig2$bm,
    ncol = 1, labels = c("A", "B", "C"),
    hjust = -5.5, vjust = 1.75
    ),
  plot_grid(plotlist = fig2$rich,
    ncol = 1, labels = c("D", "E", "F"),
    hjust = -2, vjust = 1.75
    ),
  ncol = 2,
  rel_widths = c(1, 1)
)
```

@@

We found that most temporal trends of food-webs were classified as
stable for species richness and community biomass. Roughly the same
number of stations displayed temporal decrease and increase for species
richness, and twice more stations displayed decrease in community biomass than
increase. This result is in accordance with previous meta analysis on species
richness trends of vertebrate in marine, freshwater and terrestrial systems
[@dornelas_assemblage_2014; @blowes_geography_2019; @klink_meta-analysis_2020].

@@

The temporal trends of connectance were positively linked to that of community
biomass, meaning that when connectance increased through time when community
biomass increased through time in a given site. Reversely, connectance decreased
in food-webs when community biomass decreased through time.  Indeed, food-webs
have a highly nested structure, in a way that feeding trophic links of species
at lower trophic level are a subset of the ones of species found at higher
trophic level [@kondoh_food_2010; @nordstrom_nestedness_2015]. It is because
feeding interactions are largely driven by body size, consumers being larger
than their preys [@brose_allometric_2006; @woodward_body_2005]. We also found
that community biomass losses through time were associated to a temporal
decrease of average trophic level and that community biomass temporal gains were
associated to temporal increase of average trophic level. Together with connectance, it suggests that
community biomass losses and gains are associated to losses and gains in large
bodied piscivores, i.e. biomass at the higher trophic level. Another piece of evidence
is that evenness of biomass, measured by the Pielou index, decreased through
time as community biomass increased, meaning that biomass distribution across
species becomes less even when community biomass increased. Reversely, in
food-webs that displayed temporal decrease in community biomass, we found that
evenness of biomass increased, i.e. that biomass distribution across species
becomes more even. Because predator-prey interactions relate on bodysize for a
large part [@brose_allometric_2006; @woodward_body_2005] and thus bodymass to a
certain extent, we can expect that a given gain in predators of higher trophic
level leads to a decrease of biomass evenness and reversely. In the same line,
we found that the temporal trends of community biomass were positively
associated to the temporal trends of the number and proportion of piscivorous
trophic species present in fish communities (Fig. SXX). The close association
between the temporal trends of community biomass from one side, and connectance,
average trophic level and biomass evenness from the other side, strongly suggest
that temporal trends in community biomass in natural freshwater fish communities
are associated to variations in the biomass of the top trophic level.

In contrast, we found that temporal trends of connectance and average trophic
level were not related to the temporal trends of species richness, in
contradiction with previous observation in space [@dunne_network_2006;
@danet_species_nodate]. However, we found that temporal trends of the number of
piscivorous species and piscivorous trophic species were positively associated
to the temporal trends of species richness (Fig. S3). It suggests that the
temporal decrease of species richness observed in our data are concomitant to
losses of piscivorous and non-piscivorous species, leading to the absence of
temporal trends in connectance and average trophic level and symmetrically for
the temporal increase of species richness. Furthermore, we found that the
temporal trends of biomass evenness had a positive relationship with the one of
species richness, in contradiction of what we found for temporal trends of
community biomass. Together with the fact that species richness temporal trends
had a positive relationship with the temporal trends of community biomass, it
suggests that positive and negative  temporal trends of species richness are
linked to gains and losses of redundancy in food-webs, without overall effects
on connectance and average trophic level. This is in line with the insurance
hypothesis [@yachi_biodiversity_1999] which predicts the temporal loss and gain
of species richness represent respectively loss and gain of redundancy in the
community, in the food-web structure in the present study.

The discrepancy between the effects of the temporal trends of species richness
and community biomass on food-web structure might be coined in the concepts of
trophic complementarity and redundancy [@poisot_trophic_2013;
@nordstrom_nestedness_2015]. The variations of community
biomass is related to variations in the presence of large bodied predators,
which affect both trophic complementarity, because they feed on bigger preys,
and redundancy, because their diet include the diet of lower trophic level.
Species richness temporal variation seems to affect mainly trophic redundancy
and not trophic complementarity, as average trophic level and connectance does
not change despite concomitant variations in the number of piscivorous species
(**Interesting ?**).

As studies reported community biomass or species richness temporal trends
[@hallmann_more_2017; @ceballos_biological_2017; @dornelas_assemblage_2014;
@blowes_geography_2019], we document that temporal trends of species richness
and community biomass might have different consequences on the structure of the
community. In particular, the decrease of the average trophic level associated
to the loss of community biomass could have cascading effect on ecosystem
functioning as the highest predators are predicted to increase productivity and
stabilise communities [@wang_biodiversity_2018; @shanafelt_stability_2018;
@danet_species_nodate]. In contrasts, the loss of species richness might be
linked to loss of redundancy in the food-web, highlighting that those
communities may become more at risk regarding future perturbations
[@yachi_biodiversity_1999]. Interestingly, the hypothesis that temporal trends
in community biomass is related to higher trophic level in our system has a
strong reasoning with the effect of human fisheries in marine systems which
affect primarily fish species at top trophic level [@sprules_surfing_2016].
Overall, our study highlights that monitoring several facets of biodiversity is
essential to infer how ecological communities are reshaped by loss and gain of
biomass and species richness.

@@

Surprisingly, we found that the temporal trends of connectance were not related
to the one of species richness as found spatially in most empirical food-webs
[@dunne_network_2006] and also here. We explained this pattern by the fact that
the temporal variation of species richness in our dataset seems to be linked to
temporal variation of redundancy level in the structure of food-webs rather than
overall change of their vertical structure. The same rational can applied with
average trophic level. This result is in line with other studies which found
mismatch between spatial and temporal patterns [@adler_contrasting_2007;
@johnson_testing_2008]. This discrepancy in space and time might be due to the
crucial assumption that variations in space reflect differences in assembly
stages of ecological communities and then mirrors their temporal
assembly/disassembly trajectories. Temporal variation of species richness might
relate on transitory dynamics while spatial gradient reflect aggregated dynamics
over longer time scale including a far number of processes including high
species turnover along spatial gradient. However, we fount that the positive
relationship between evenness of biomass and species richness were consistent in
space and time. Another reason might be that the spatial gradient displays a
higher range of variation for species richness than the temporal stations (Fig.
S2). However, we found strong relationships between the temporal trend of
species richness from one side and the temporal trends of community biomass,
biomass evenness, and the number of piscivorous species from the other side (Fig. S3).
